---
output: pdf_document
---

<br>

# **Methylation Statistics**

<br>

This Analysis can check the basic stats about the methylation data such as coverage and percent methylation.   

<br>

```{r}
## read bam bismark file and make meth object

library(methylKit)

destfile="methylKit.RDS"

if( ! file.exists(destfile) ){
    source("/home/adminrig/src/short_read_assembly/bin/R/MethylKit.makeRDS.R")
}

# files<-as.list( list.files(path = "DMR/BAM", pattern = "sort.bam$", recursive = FALSE, full.names = TRUE) )
# samples<-as.list( gsub( ".+\\/(.+)_\\w{6,8}_L.+", "\\1", files ) )
# 
# objs<-processBismarkAln(location=files, sample.id=samples, assembly="hg19", save.folder=NULL, save.context=NULL, read.context="CpG", nolap=FALSE, mincov=10, minqual=20, phred64=FALSE, treatment=rep(0, length(samples)))
# 
# saveRDS(objs, "methylKit.RDS")


objs <- readRDS("methylKit.RDS")

myobj<-objs
meth=methylKit::unite(myobj, destrand=FALSE)

sample_ids <- getSampleID(meth)

getTreatment(meth) <- 1:length(sample_ids)

#help(package="methylKit")
#load("/home/adminrig/workspace.min/MethylSeq/KNIH.IncurableDisease/DMR/BAM/meth.RDS")

```


<br>

## Histogram of % CpG methylation

<br>

The following result show the histogram for percent methylation distribution. The figure below is the histogram and numbers on bars denote what percentage of locations are contained in that bin.
Typically, percent methylation histogram should have two peaks on both ends. In any given cell, any given base are either methylated or not. Therefore, looking at many cells should yield a similar pattern where we see lots of locations with high methylation and lots of locations with low methylation.   

<br>
<br>
<br>
<br>
<br>

<!-- ```{r, fig.width=12, fig.height=12} -->
```{r}

n_col = 3
n_row = ceiling ( length( sample_ids ) / n_col )


# par(mfrow=c(n_row,n_col))

par(mfrow=c(2,1))
for( i in 1:length(sample_ids) )
{
  #print(samples[i])
  
  #getMethylationStats( myobj[[i]], plot=TRUE, both.strands=FALSE)
  A<-myobj[[i]]
  
  h = hist( (A@.Data[[6]] )/( A@.Data[[5]] ) * 100, plot=FALSE, breaks = 10)
  h$density = round( h$counts/sum(h$counts)*100, 1)
  plot(h,freq=FALSE, label=TRUE, main= A@sample.id, col="cornflowerblue", xlab= "% methylation per base", ylab="Percent (%)")

}



```

\clearpage

<br>

## Histogram of % CpG Coverage

<br>

The following result can also plot the read coverage per base information in a similar way, again numbers on bars denote what percentage of locations are contained in that bin.
Experiments that are highly suffering from PCR duplication bias will have a secondary peak towards the right hand side of the histogram.   

<br>


```{r, fig.width=12, fig.height=12}


# par(mfrow=c(n_row,n_col))
par(mfrow=c(2,1))
#par(mfg=c(1,1))
# layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow = TRUE))

for( i in 1:length(sample_ids) )

{
  #print(samples[i])
  
  #getMethylationStats( myobj[[i]], plot=TRUE, both.strands=FALSE)
  A<-myobj[[i]]
  
  #getCoverageStats(A, plot=TRUE, both.strands = FALSE, label =FALSE, breaks = 20, main="")
  h = hist(  log10(  A@.Data[[5]]  ) , plot=FALSE, breaks = 20)
  h$density = round( h$counts/sum(h$counts)*100, 1)
  plot(h,freq=FALSE, label=TRUE, main= A@sample.id, col="chartreuse4", xlab= "log10 of read coverage per base", ylab="Percent (%)")

}


#plot.new()

```

\clearpage



<br>

# **Comparative Analysis**

<br>


<br>

## Sample Correlation

<br>

This analysis can check the correlation between samples using methylation data and plot scatter plot.  

<br>

```{r, fig.width=12, fig.height=12}
#pdf("Correlation.pdf", width=8, height=9)
aa<- capture.output( getCorrelation(meth, plot=TRUE) )
#dev.off()

```

\clearpage


<!-- <br> -->

<!-- ## Clustering Samples -->

<!-- <br> -->

<!-- ```{r, warning=FALSE} -->
<!-- #pdf("HC.pdf", width=5, height=5 ) -->
<!-- hc<-clusterSamples(meth, dist="correlation", method="ward", plot=TRUE) -->
<!-- #dev.off() -->

<!-- ``` -->

<!-- \clearpage -->

<br>

## CpG methylation PCA

<br>

PCA analysis plot PC1 and PC2 axis and a scatter plot of your samples on those asix which will reveal how they cluster.   
 
<br> 

```{r, include=FALSE}

#pdf("PCA.v1.pdf", width=5, height=5 )
#PCASamples(meth)
#dev.off()

library(ggfortify)
library(ggplot2)
library(ggrepel)

#pdf("tmp.pdf", width=5, height=5 )
PCA_out<-PCASamples(meth,obj.return=TRUE)
#dev.off()

df_out_r <- as.data.frame( PCA_out$x )
df_out_r$feature <- row.names(df_out_r)

```


```{r}
# http://rfriend.tistory.com/61
# print(PCA_out)

PC1_Proportion <- round( (summary(PCA_out)$importance[2, ])[1]*100, 2)
PC2_Proportion <- round( (summary(PCA_out)$importance[2, ])[2]*100, 2)
```



```{r, fig.width=6, fig.height=6}
#pdf("PCA.v2.pdf", width=4, height=4)
ggplot(df_out_r,aes(x=PC1,y=PC2,label=feature,color=feature )) +
  theme_bw() +
  #theme_classic() +
  geom_point() +
  geom_text_repel(aes(label=feature), size=3) +
  theme(legend.position="none") + 
  labs(x = paste0("PC1 (", PC1_Proportion, "% explained var.)"), y = paste0("PC2 (", PC2_Proportion, "% explained var.)"))




#dev.off()


```


\clearpage

<!-- <div style="width:1000px; height:1000px"> -->
<!-- ![PCA](/home/adminrig/workspace.min/MethylSeq/KNIH.IncurableDisease/DMR/BAM/PCA.v2.pdf){} -->
<!-- </div> -->




```{r}
# head( objs )
# head( myobj )

# objs<-processBismarkAln(location=files, sample.id=samples, assembly="hg19", save.folder=NULL, save.context=NULL, read.context="CpG", nolap=FALSE, mincov=10, minqual=20, phred64=FALSE, treatment=rep(0, length(samples)))
# 
# 
# 
# myobj<-objs
# meth=unite(myobj, destrand=FALSE)
# getTreatment(meth) <- 1:length(samples)

```



<!-- <br> -->

<!-- # DMR Analysis Result  -->

<!-- <br> -->

<!-- ```{r, include=FALSE} -->
<!-- df<-read_tsv("DMRs") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #head(df) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>%  -->
<!--   group_by(Analysis) %>%  -->
<!--   summarise(Count=n()) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>%  -->
<!--   group_by(Analysis, Region) %>%  -->
<!--   summarise(Count=n()) %>%  -->
<!--   spread( Region, Count) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- breaks <-c(0,1,5,10,20,30,50,100,200,300,500,5000) -->

<!-- width_table<- -->
<!--   table( list(df$Analysis, cut(df$Width, breaks=breaks)) ) %>%  -->
<!--   as.data.frame() -->

<!-- colnames(width_table)<-c("Analysis", "Range", "Count") -->

<!-- dplyr::select(width_table, Range, Analysis, Count) %>%  -->
<!--   spread(Analysis, Count) -->


<!-- ``` -->

<!-- ```{r, fig.width=9, fig.height=8} -->
<!-- ggplot(width_table,  aes(Range, Count, fill=Analysis)) + geom_col() + facet_grid(Analysis~.) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- head(df) -->
<!-- ``` -->













